---
title: "hw_3_jw4007"
output: github_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, message = FALSE)
```

```{r load_libraries}
library(tidyverse)
library(readxl)
```

# problem 1

```{r}
library(p8105.datasets)
data("instacart")
```
### How many aisles are there, and which aisles are the most items ordered from?
```{r}
instacart %>%
  group_by(aisle) %>%
  count(sort = TRUE)
```
There are 134 aisles and the aisles with the most items ordered is fresh vegetables.

### Make a plot that shows the number of items ordered in each aisle, limiting this to aisles with more than 10000 items ordered. Arrange aisles sensibly, and organize your plot so others can read it.
```{r}
aisle_1000 =
  count(instacart, aisle) %>%
  filter(n > 10000) %>%
  mutate(aisle = forcats::fct_reorder(aisle, n))

ggplot(aisle_1000, aes(x = aisle, y = n)) +
  geom_col() +
  labs(title = "The number of items ordered in each aisle",
       x = "aisle",
       y = "the number of items ordered") +
  theme(
    plot.title = element_text(size = 16),
    axis.title.x = element_text(size = 10),
    axis.title.y = element_text(size = 10),
    axis.text.y = element_text(size = 8)
  ) +
  coord_flip()
  
```
### Make a table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”. Include the number of times each item is ordered in your table.
```{r}
popular_aisles = 
  group_by(instacart, aisle) %>%
  select()
```

# problem 2

### Load, tidy, and create "weekday vs weekend" variable. 
```{r}
cardio = read.csv("./accel_data.csv") %>%
  janitor::clean_names() %>%
  pivot_longer(
    activity_1:activity_1440,
    names_to = "minutes",
    values_to = "activity_counts",
    names_prefix = "activity_"
  ) %>%
  mutate(weekday_vs_weekend = 
           ifelse(day == "Saturday", "weekend",
                  ifelse(day == "Sunday", "weekend", "weekday"))) %>%
  mutate(minutes = as.numeric(minutes))
```
There is `r nrow(cardio)` observations and `r ncol(cardio)` variables in the dataset. The variables are `r names(cardio)`

### Aggregate across minutes to create a total activity variable for each day, and create a table showing these totals. Are any trends apparent?
```{r}
daily_cardio = cardio %>%
  group_by(week, day) %>%
  summarise(daily_activity = sum(activity_counts), .groups = "drop") %>%
  pivot_wider(names_from = day, values_from = daily_activity) %>%
  select("week","Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday")

knitr::kable(daily_cardio)
```
It seems that the daily activity on weekdays are less than weekends.

### Make a single-panel plot that shows the 24-hour activity time courses for each day and use color to indicate day of the week.
```{r}
ggplot(cardio, 
       aes(x = minutes, y = activity_counts, color = day)) +
  geom_line() +
  labs(title = "24-hour avtivity time course",
       x = "Time(minutes)",
       y = "Avtivity") +
  scale_x_continuous(
    breaks = c(0, 360, 720, 1080, 1440),
    labels = c("12a.m.", "6a.m.", "12p.m.", "6p.m.", "11:59p.m."),
    limits = c(0, 1440))
  
```

```{r}

```

```{r}

```

